On Math

### 2025-04-15 cards/paper.tex
derivatives and integrals

@@

@@

>>>>

\section{Geometry}
\subsection{Rigid motions}
Or sometimes - any transformations that does not change the distances
and angles between points in a figure. They are ANGLE-PRESERVING and
DISTANCE-PRESERVING.

Rigid motions types:
- translations
- rotations
- reflections

Not a rigid motions:
- dilations
- stretches

@@

\subsection{Factor theorem}
For polynomial, if we know root, we can determine a factor of the
polynomial. If $p(x)$ is a polynomial and $x=r$ is a root of $p(x)$ then
\begin{equation}
  \begin{gathered}
  p(r) = 0 \\
  so (x-r) is a factor of p(x) \\
  generally: (ax-b) is a factor of a polynomial p(x) then \\
  x = \frac{b}{a} is a root of p(x)
\end{gathered}
\end{equation}

\subsection{Inverse functions}
Inverse functions are denoted by $f^{-1}(x)$.

For $f(x)$ that maps $x=2$ to $y=0$ an inverse function $f^{-1}(x)$ instantiation would be
$f^{-1}(0)=0$.

Generally it is a simple composition:
\begin{equation}
  \begin{gathered}
    f^{-1}(f(x)) = x \\
    so \\
    (f^{-1} \circ f)(x) = x
  \end{gathered}
\end{equation}

Calculate inverse function
1. swap $x$ and $y$
2. solve for $y$

Example:
\begin{equation}
  \begin{gathered}
    f(x) = 3x-9 \\
    y = 3x-9 \\
    x = 3y-9 \\
    x-3y=-9 \\
    x+9=3y \\
    \frac{x+9}{3}=y \\
    f^{-1}(x)=\frac{x+9}{3}
  \end{gathered}
\end{equation}

Domain and range of inverse function:
Domain of a function $f(x)$ is the range of its inverse function $f^{-1}(x)$.
Range of a function $f(x)$ is the domain of its inverse function $f^{-1}(x)$.

\subsection{Multiplicities of the roots of a polynomial}
\begin{equation}
  p(x) = (x-5)^2(x+8)
\end{equation}
for that polynomial roots are $x=5$ and $x=-8$ but factor $(x-5)^2$
appears TWICE so it is a MULTIPLE ROOT
$(x+8)$ on the other hand is a SIMPLE ROOT.

@@

\section{Derivatives}
Average rate of change of the function $f(x)$ on the interval $[a, a+h]$

\begin{equation}
  \frac{\Delta y}{\Delta x} = \frac{f(a+h) - f(a)}{h}
\end{equation}

Instanteous rate of change of the function at the point $x=a$ is the limit of the average
rate of change as $h \rightarrow 0$. It's denoted by $f'(a)$ and given by:

\begin{equation}
  f'(a) = \lim_{h \rightarrow 0}\frac{f(a+h) - f(a)}{h}
\end{equation}

and its value is equal to the slope of the tangent line to the curve $y = f(x)$ at $x=a$.
Also it can be defined as:
\begin{equation}
  f'(a) = \lim_{x \rightarrow a}\frac{f(x) - f(a)}{x-a}
\end{equation}

\section{Integrals}
Also called \textit{antiderivatives}. To represent integrals

Integral of a sum of functions
\begin{equation}
  \int f(x) \pm g(x) \, dx = \int f(x) \, dx \pm \int g(x) \, dx
\end{equation}

Integral of a reciprocal function
\begin{equation}
  \int \frac{1}{x} \, dx = \int \ln abs(x) + C
\end{equation}
and also:
\begin{equation}
  \int \frac{1}{x} \, dx = \int \ln|x| + C = \ln abs(x) + \ln K = \ln (K (abs(x))
\end{equation}

Integral of an exponential function
\begin{equation}
  \int e^x \, dx = e^x + C
\end{equation}
with general base we know that
\begin{equation}
  \frac{d}{dx} (a^x) = a^x \ln a
\end{equation}
so reversing:
\begin{equation}
  \int a^x \, dx = \frac{a^x}{\ln a} + C
\end{equation}

Marked integral for future
\begin{equation}
  \int_{a}^{b} f(x) \, dx
\end{equation}
